
% !TEX encoding = UTF-8 Unicode 
% !TEX root = FieldGuide.tex


\clearpage

%\thispagestyle{}
%~
%\vspace{18em}
%~
\SecS{Acknowledgments}
\phantomsection\addcontentsline{toc}{section}{Acknowledgments \& Version History} 
%\begin{center}
%\textsc{Acknowledgments}
%\end{center}


In curating this collection of distributions, I have benefited greatly from Johnson, Kotz, and Balakrishnan's monumental compendiums~\cite{Johnson1994, Johnson1995}, Eric Weisstein's MathWorld, the Leemis chart of Univariate Distribution Relationships~\cite{Leemis2008, Leemis2012}, and myriad pseudo-anonymous contributors to Wikipedia. Additional contributions are noted in the version history below.

% Add Devroye1986 ?


\begin{center}
\textsc{Version History}
\end{center}

{
\small \tightstretch

0.13 (2019-??-??)
Added
log-Cauchy \eqref{LogCauchy}


~

0.12 (2019-02-14) 
Added
Epanechnikov \eqref{Epanechnikov}, 
Libby-Novick \eqref{LibbyNovick},
Gauss hypergeometric \eqref{GaussHypergeometric},
and
confluent hypergeometric \eqref{Confluent}
distributions. 
% Fixed many typos
% Merged Pearson III and gamma distributions.
% Added figure for log-gamma
% Renamed Prentice to beta-logistic
% Add source control: https://github.com/gecrooks/fieldguide

~

0.11 (2017-06-19)
\sloppy{ Added hyperbola \eqref{Hyperbola}, Halphen \eqref{Halphen}, Halphen B \eqref{HalphenB}, inverse Halphen B \eqref{InvHalphenB}, generalized Halphen \eqref{GenHalphen}, Sichel \eqref{Sichel} and Appell Beta~\eqref{AppellBeta} distributions. Thanks to Saralees Nadarajah.}
% Kudos to Nadarajah for AppellBeta paper.

~

0.10 (2017-02-08)
Added K \eqref{K}  and generalized K  \eqref{GenK} distributions.
% Fixed typos.
% Additional endorsements.
Clarified notation and nomenclature.
Thanks to Harish Vangala.

~

0.9 (2016-10-18)
Added 
	pseudo Voigt \eqref{PseudoVoigt}, 
and
	Student's $t_3$ \eqref{StudentsT3}
distributions.
Reparameterized hyperbolic sine \eqref{HyperbolicSine} distribution.
Renamed inverse Burr to Dagum \eqref{Dagum}. %, and added location scale
Derived limit of unit gamma to log-normal (p\pageref{UnitGammaToLogNormal}). 
Corrected spelling of  ``arrises'' (sharp edges formed by the meeting of surfaces) to ``arises'' (emerge; become apparent). 
% Added figure of limits and reworked limits appendix
% Added figure for order statistics appendix
% Added Endorsements page
% Moved Pearson Exponential to Notes until figure out best organization.
% Removed stretched exponential distribution (Redundant with Weibull)
% Removed Moffat, unclear if correct definition, or if redundant.
% 50% proofed equations

~

0.8 (2016-08-30)
The Unprincipled edition:
Added Moyal distribution \eqref{Moyal}, a special case of the gamma-exponential distribution.
Corrected spelling of ``principle'' to ``principal''. Thanks to Matthew Hankins and Mara Averick.
% Fixed typos and improved formatting.
% Added source control
% Renamed mixture distributions to compound distributions. 
% Added exponential twist, but not done yet.
% Changed all distribution names to links.
% Fixed overfull hboxes

~

0.7 (2016-04-05) 
%
Added Hohlfeld distribution. %and Gauss hypergeometric distributions.
Added appendix on limits.
Reformatted and rationalized distribution hierarchy diagrams. 
%Fixed typos and improved formatting.
Thanks to Phill Geissler.

~

0.6 (2014-12-22)
%
Total of 147 named simple, unimodal, univariate, continuous probability distributions, and at least as many synonyms.
Added appendix on the algebra of random variables.
Added Box-Muller transformation.
For the gamma-exponential distribution, switched the sign on the parameter $\alpha$.
Fixed the relation between beta distributions and ratios of gamma distributions ($\alpha$ and $\gamma$ were switched in most cases).
%Fixed typos and improved formatting.
Thanks to Fabian Kr\"uger and Lawrence Leemis.


~


0.5 (2013-07-01)
Added uniform product, half generalized Pearson VII, half exponential power distributions, GUD and q-Type distributions. 
Moved Pearson IV to own section.
Fixed errors in Inverse Gaussian.
Added random variate generation to appendix.
% Fixed typos. 
Thanks to David Sivak, Dieter Grientschnig, Srividya Iyer-Biswas, and Shervin Fatehi.

~

 0.4 (2012-03-01) 
% Fixed typos.
Added erratics. 
Moved gamma distribution to own section.
Renamed log-gamma to gamma-exponential.
Added permalink.
Added new tree of distributions. 
Thanks to David Sivak and Frederik Beaujean.

~

0.3 (2011-06-40) 
%Fixed many typos.
Added tree of distributions. 
%Improved formatting. 

~

0.2 (2011-03-01) 
%Fixed many typos.
Expanded families.  
Thanks to David Sivak. 

~

 0.1 (2011-01-16)
 Initial release. Organize over 100  simple, continuous, univariate probability distributions into 14 families. Greatly expands on previous paper that discussed the Amoroso and gamma-exponential families~\cite{_amoroso}. Thanks to David Sivak, Edward E. Ayoub, and Francis J. O'Brien.

%~
%0.0 (c2010) Genesis, during a particularly enthralling session of the Berkeley Statistical Mechanics meeting. 
}



\normalstretch
\newpage ~\newpage

\SSecS{Endorsements}
~\\

\noindent
``\emph{Ridiculously useful}'' -- Mara Averick\footnote{\url{https://twitter.com/dataandme/status/770732084872810496}}
\\

\noindent
``\emph{I can't stress how useful I've found this. I wish I'd had a printout of it by my desk every day for the last 6 years}''-- Guillermo Roditi Dominguez\footnote{\url{https://twitter.com/groditi/status/772266190190194688}}
\\

\noindent
``\emph{Abramowitz and Stegun for probability distributions}''-- Kranthi K. Mandadapu\footnote{Thursday Lunch with Scientists}
\\

\noindent
``\emph{I had no idea how much I needed this guide.}''-- Daniel J. Harris\footnote{\url{https://twitter.com/DHarrisPsyc/status/870614354529370112}}
\\



\noindent
``\emph{Who are you? How did you get in my house?}'' -- Donald Knuth\footnote{\url{https://xkcd.com/163/}}
\\

